---
title: "Final Project Report"
author: "AUTHOR NAME(S)"
date: "Due: DATE"
output: html_notebook
---

# Set Up

#### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, `lubridate`, and all others used
library(mosaic)
library(tidyverse)
library(tidyr)
library(dcData)

library(dplyr)
library(ggplot2)

library(party)
library(lubridate)
library(ggplot2)
library(dplyr)
library(esquisse)
library(leaflet)
library(rvest)
library(olsrr)
```

## Research Questions

- Which Airports experience more airplane accidents compared to the majority and what factors are usually significantly similar     between the incidents?


## Data Access
```{r}
#Data intake
Transportation <- read.file("AviationData.csv") 
AirPorts <- read.file("ntsb_aviation_data.csv")
```



```{r}
# Main Data inspection
summary(Transportation)
glimpse(Transportation)
head(Transportation)
nrow(Transportation)
```


```{r}
# Other Data inspection
glimpse(AirPorts)
head(AirPorts)
nrow(AirPorts)
names(AirPorts)
```




## Data Wrangling

(+) general data wrangling using various data verbs like filter, mutate, summarise, arrange, group_by, etc.
(+) joins for multiple data tables. 
(+) pivot_wider & pivot_longer--or similar--to stack/unstack variables
(+) regular expressions
(+) reduction and/or transformation functions like mean, sum, max, min, n(), rank, pmin, etc. 
(+) user-defined functions 
(+) loops/control flow 

```{r}
Transportation %>%
  filter(Injury.Severity == "Fatal") %>%
  head(20)
```
```{r}
 Joint_Transport %>%
    mutate(AmateurORNot = ifelse(Amateur.Built == 'Yes', 1, 0))
```


```{r}

  
  Joint_Transport <-
  Transportation %>%
  select(Investigation.Type, Accident.Number, Injury.Severity, Aircraft.damage, Number.of.Engines, Total.Fatal.Injuries, Total.Serious.Injuries) %>%
  left_join(AirPorts %>% select(NTSB_RPRT_NBR, FLIGHT_PHASE_DESC ), by = c("Accident.Number" = "NTSB_RPRT_NBR")) %>%
  na.omit(Joint_Transport) 



Joint_Transport
```






## loop + user define function##
```{r}
set.seed(184)
Latitude = Transportation$Latitude
longitude= Transportation$Longitude
Total.Fatal.Injuries =Transportation$Total.Fatal.Injuries
## define the variable
model1 <- glm(Total.Fatal.Injuries ~ Latitude+longitude)
## define the model for forecasting the fatal injuries
b1 <- coef(model1)[[1]]
b2 <- coef(model1)[[2]]#Latitude
b3 <- coef(model1)[[3]]#longitude
## create a vector space for storing the parameter of data
summary(model1)
# check the parameter
y= Total.Fatal.Injuries
par(mfrow= c(1,2))
hist(Latitude)
hist(longitude)
## check the disatribution and spread of the variable
latimean <- mean(Latitude, na.rm = TRUE)
logimean <- mean(longitude, na.rm = TRUE)
latisd <- sd(Latitude, na.rm = TRUE)
logisd <- sd(longitude, na.rm = TRUE)
## calculate the parameter for looping
library(data.table)
n = 1000
itera = 1000
## runtime
store1 <-rep(0,itera)
store2 <-rep(0,itera)
strore_y <-rep(0,itera)
## store the data for bi (empty space)
x<- mean(Total.Fatal.Injuries)
for (i in 1:itera){
  error = rnorm(n, mean = 1, sd = 1) 
  xlatitude = rnorm(n, mean = latimean, sd = latisd)
  xlogitude = rnorm(n, mean =logimean, sd = logisd)
  
  Y_i = b1 + b2*xlatitude+b3*xlogitude+error
  
  fun_i <-glm(Y_i~xlatitude+xlogitude)
  
  store1[i] <- coef(fun_i)[[2]]
  store2[i] <-coef(fun_i)[[3]]
  strore_y[i] <-Y_i[i]
  
  data_i = data.table(Y = Y_i)
}
## run  the model with loop
sd1 <-sd(strore_y)
mean1 <- mean(strore_y)
## get the parameter
plot(density(strore_y), main = "Density plot of Fatal injuries predicted", ylim = c(0,0.4), xlim = c(0,8))
curve(dnorm(x, mean1, sd1), col="red", add=TRUE)
hist( strore_y, prob=TRUE, main = "histogram of Fatal injuries predicted", ylim = c(0,0.4), xlim = c(0,8))
curve(dnorm(x, mean1, sd1), col="red", add=TRUE)
## check the predicted is valid enough
alpha<-0.05
quantile(strore_y,probs=c(alpha/2,1-alpha/2))
## predict injury numbers result
```


## Data Visualization 

(*1*) relevant to stated research question for the analysis
(*2*) are neat with professional appearance including titles, axis labels, guides, etc 
(*3*) include at least one effective display of many--at least 3--variables (cannot use a map or decision tree for this requirement)
(*4*) include 3 of the following 5 visualization techniques learned in STAT 184:
(+) use of multiple geoms such as points, density, lines, boxplots, etc 
(+) use of multiple aesthetics--not necessarily all in the same graph--such as color, size, shape, x/y position, facets, alpha, etc 
(+) layered graphics such as points and accompanying smoother, points and accompanying boxplots, overlaid density distributions, etc 
(+) leaflet or choropleth maps 
(+) decision tree and/or dendogram displaying exploratory machine learning results

```{r}
library(dplyr)
library(ggplot2)

Transportation %>%
 filter(!(Investigation.Type %in% "")) %>%
 filter(!(Aircraft.damage %in% 
    "")) %>%
 filter(!(Aircraft.Category %in% "")) %>%
 filter(!(Amateur.Built %in% 
    "")) %>%
 filter(!(Engine.Type %in% "")) %>%
 filter(!(FAR.Description %in% "")) %>%
 
    filter(!(Schedule %in% "")) %>%
 filter(!(Purpose.of.flight %in% "")) %>%
 filter(!(Weather.Condition %in% 
    "")) %>%
 filter(!(Broad.phase.of.flight %in% "")) %>%
 filter(!(Report.Status %in% 
    "")) %>%
 ggplot() +
 aes(x = Engine.Type, weight = Total.Fatal.Injuries) +
 geom_bar(fill = "#0c4c8a") +
 theme_minimal()
```


```{r}
Transportation %>%
 filter(!(Investigation.Type %in% "")) %>%
 filter(!(Aircraft.damage %in% 
    "")) %>%
 filter(!(Aircraft.Category %in% "")) %>%
 filter(!(Amateur.Built %in% 
    "")) %>%
 filter(!(Engine.Type %in% "")) %>%
 filter(!(FAR.Description %in% "")) %>%
 
    filter(!(Schedule %in% "")) %>%
 filter(!(Purpose.of.flight %in% "")) %>%
 filter(!(Weather.Condition %in% 
    "")) %>%
 filter(!(Broad.phase.of.flight %in% "")) %>%
 filter(!(Report.Status %in% 
    "")) %>%
 ggplot() +
 aes(x = Weather.Condition, y = Total.Fatal.Injuries, colour = Aircraft.Category) +
 geom_boxplot(fill = "#0c4c8a") +
 scale_color_brewer(palette = "YlGnBu") +
 theme_minimal() +
 ylim(0L, 20L)
```



```{r}
  Danger <-
  Joint_Transport1 %>%
  ctree( Aircraft.damage == "Substantial" ~ Number.of.Engines + Total.Serious.Injuries  , data = .)

plot(Danger, type ="simple")
```




### Purpose of investigating this topic




### Problems faced during this project 






